---
title: "Creating new actions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating new actions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(scenes)
```

Actions (class `scene_action`) contain the functions that are used to process a `request` and determine which `ui` should be displayed.
We provide five actions:

- `req_has_cookie()`
- `req_has_query()`
- `req_uses_method()`
- `req_uses_get()`
- `req_uses_get()`

A scene can require multiple actions to be true.
If you need to rely on multiple query parameters, or a cookie and a query parameter, stringing together multiple actions will probably suffice.

However, you may wish to construct more complicated actions, such as multiple alternative query parameters (`or`, not `and`), or check a request parameter that we do not support.
You can do so using `construct_action()`.

## Check function

Underlying each action is a **check function,** a function that takes a request (and potentially other arguments), and returns `TRUE` or `FALSE`.

Here we'll implement a check for the language preferred by the user, which is sent in the `HTTP_ACCEPT_LANGUAGE` property of the request.

We include `_impl` in the name of this function to specify that it's the implentation function, as opposed to the main wrapper that we'll create below.
For this we'll just look for the supplied "language" inside the `HTTP_ACCEPT_LANGUAGE` object.
In an exported function, we'd probably more carefully parse that object.

```{r req_accepts_language}
req_accepts_language_impl <- function(request, language) {
  stringr::str_detect(
    tolower(request$HTTP_ACCEPT_LANGUAGE), 
    tolower(language)
  )
}
req_accepts_language_impl(
  list(HTTP_ACCEPT_LANGUAGE = "en-US,en;es-MX,es;fr-CA,fr"), 
  "fr"
)
```

## Methods

Almost all actions will expect the `GET` method.
However, it is possible for shiny apps to respond to requests using other HTTP methods.
If your action should work with a different HTTP method, specify that in the call to the contructor.

## Construct action

Construct the action using the `construct_action()` function.

```{r construct_action}
req_accepts_language <- function(language) {
  construct_action(
    fn = req_accepts_language_impl,
    language = language,
    # We're using the defaults for these arguments, but I'll specify them for
    # clarity.
    negate = FALSE,
    methods = "GET"
  )
}
```

Now you can use this action to construct scenes, just like any other action.
